\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-06-11 bmc_section.tex
Add the explicit enumeration of paths.

We describe the the procedure $\mathsf{BoundedModelCheck}$ used at
Line~\ref{algoline:bmc} in Algorithm~\ref{algo:s3camBMC}
that checks if the $\epsilon$-precision Relational Transition System
$\RTSAbstraction{\epsilon}{\System}$ violates the safety property
$\psi(\vx)$ in at most $k$ steps.

We describe two possible implementations:
(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic} encoding of
the paths of $\RTSAbstraction{\epsilon}{\System}$ of length $k$ that 
violates $\psi$, and (2) an algorithm that explicitly enumerates the
``discrete'' paths formed cells $\Cells$ and the transition
$\CellEdges$, and then checks if the relations of the variables $\vx$
on the visited edges hold for for each path (this is in similar in
spirit to~\cite{DBLP:conf/fmcad/BuLWL08}).

%% BMC macro
\newcommand{\bvars}[1]{\ensuremath{{B}^{#1}}}
\newcommand{\rvars}[1]{\ensuremath{{\vx}^{#1}}}
\newcommand{\vars}[1]{\ensuremath{{V}^{#1}}}
\newcommand{\cellvar}[1]{\ensuremath{{cell}^{#1}}}
\newcommand{\initbmc}{\ensuremath{INIT}}
\newcommand{\encbmc}{\ensuremath{BMC}}
\newcommand{\errorbmc}{\ensuremath{ERROR}}
\newcommand{\transbmc}{\ensuremath{TRANS}}
\newcommand{\edgebmc}{\ensuremath{EDGE}}
\newcommand{\subs}[3]{\ensuremath{{#1}[{#2}/{#3}]}}

Given a set of variables $X$, we denote with $X^i=\{x^i | x \in X\}$
the copy of the variables $X$ renamed with the $^i$ superscript.
Given a Boolean formula $\exprs{V}$
%% $\subs{\exprs{V}}{z}{x}$ is the formula $\exprs{V}$ where all the
%% occurrences of the variable $x$ are replaced with $z$.
We use the shorthand $\subs{\exprs{V}}{X^i}{X}$ to denote the formula
where every variable $x$ is substituted to $x^+$

We encode the finite set of cells $\Cells$ with a set of Boolean
variables $\bvars{}$. For clarity, given the cell $C \in \Cells$ we
write $\cellvar{} = C$, instead of showing its encoding with $\bvars{}$.
%
In the following, we interpret each variable $x \in \vx$ as a Real
valued variable and we refer to all the Boolean and Real variables
with the set $\vars{} = \rvars{} \vee \bvars{}$
%
We encode the set of paths of length $k$ of
$\RTSAbstraction{\epsilon}{\System}$ that violates the property $\psi$
as the quantifier first-order logic formula:
\begin{eqnarray*}
\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k) & := &
\initbmc(\vars{0}) \wedge
\bigwedge_{i=1}^{k}{\transbmc(\vars{i-1}, \vars{i})} \wedge
\neg \subs{\psi}{\rvars{i}}{\rvars{}}\\
% init
\initbmc(\vars{i}) & := &
\bigwedge_{C \in \InitCells}{(\cellvar{i} = C) \wedge
\subs{\CellLabelingFunction(C)}{\rvars{i}}{\rvars{}}
}
\\
% trans
\transbmc(\vars{i}, \vars{i+1}) & := &
  \bigvee_{(C,C^+) \in \CellLabelingFunction}{
    \edgebmc_{(C,C^+)}(\vars{i}, \vars{i+1})
  } \\
\edgebmc_{(C,C^+)}(\vars{i}, \vars{i+1}) & := &
(\cellvar{i} = C) \wedge (\cellvar{i+1} = C^+) \wedge  \\
&& \subs{\CellLabelingFunction(C)}{\rvars{i}}{\rvars{}} \wedge
   \subs{\CellLabelingFunction(C^+)}{\rvars{i+1}}{\rvars{}} \wedge \\
&& \subs{\subs{\EdgeLabelingFunction((C,C^+))}{\rvars{i}}{\rvars{}}}{\rvars{i+1}}{{\rvars{+}}} \\
\end{eqnarray*}

The formula
$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k)$ is satisfiable
if and only if there exists a path of length $k$.
The procedure
$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System}, \psi, k)$
checks the satisfiability of the sequence of formulas 
$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, 0)$, 
$\ldots$,
$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k)$
with a SMT solver, returning as soon as one of the formula is
satisfiable (i.e. as soon as it found a path in
$\RTSAbstraction{\epsilon}{\System}$).

The PWA transition system is a finite location, infinite state
transition system. As all the guards and transitions are affine, a
bounded reachability query is equivalent to checking all
combinations of discrete locations, where each combination can be
summarized as a linear program. As the time is bounded, the number of
combinations are finite, but, exponential in number.

It should be noted that an explicit execution of the transition system
represents a directed acyclic graph, with each location being a node,
and a directed edge from each node to all nodes reachable in forward
time. A path on this graph is then a linear program, with each
branching node corresponding to a logical disjunction.

Temporal properties of PWA transitions systems can be checked using
off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
over infinite state transitions systems using SMT solvers.
Furthermore, lazy SMT solvers can be employed for this specific
problem instance to achieve better efficiency~\cite{shoukry2017smc}.
We now show how relational PWA abstractions can be viewed as PWA
transition systems, and how the problem of falsification can
be answered by a BMC query.

>>>>

%% BMC macro
\newcommand{\unsafebmc}{\ensuremath{{\psi}_{\UnsafeStates}}}
\newcommand{\bvars}[1]{\ensuremath{{B}^{#1}}}
\newcommand{\rvars}[1]{\ensuremath{{\vx}^{#1}}}
\newcommand{\vars}[1]{\ensuremath{{V}^{#1}}}
\newcommand{\cellvar}[1]{\ensuremath{{cell}^{#1}}}
\newcommand{\initbmc}{\ensuremath{INIT}}
\newcommand{\encbmc}{\ensuremath{BMC}}
\newcommand{\transbmc}{\ensuremath{TRANS}}
\newcommand{\edgebmc}{\ensuremath{EDGE}}
\newcommand{\subs}[3]{\ensuremath{{#1}[{#2}/{#3}]}}

\newcommand{\abspathbmc}[0]{\ensuremath{\Pi_{abs}}}
\newcommand{\concbmc}[1]{\ensuremath{CONC_{{#1}}}}
%
We describe the the procedure $\mathsf{BoundedModelCheck}$ used at
Line~\ref{algoline:bmc} in Algorithm~\ref{algo:s3camBMC}
that checks if the $\epsilon$-precision Relational Transition System
$\RTSAbstraction{\epsilon}{\System}$ reaches an error condition
$\UnsafeStates$ in at most $k$ steps. We assume that the set of unsafe
states $\UnsafeStates$ can be represented with a formula
$\unsafebmc(\vx)$

We describe two possible implementations:
(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic} encoding of
the paths of $\RTSAbstraction{\epsilon}{\System}$ of length $k$ that 
reaches $\unsafebmc$, and (2) an algorithm that explicitly enumerates the
``discrete'' paths formed cells $\Cells$ and the transition
$\CellEdges$, and then checks if the relations of the variables $\vx$
on the visited edges hold for for each path (this is in similar in
spirit to~\cite{DBLP:conf/fmcad/BuLWL08}).

\paragraph{Symbolic path exploration.}
%
%
Given a set of variables $X$, we denote with $X^i=\{x^i | x \in X\}$
the copy of the variables $X$ renamed with the $^i$ superscript.
Given a Boolean formula $\exprs{V}$
%% $\subs{\exprs{V}}{z}{x}$ is the formula $\exprs{V}$ where all the
%% occurrences of the variable $x$ are replaced with $z$.
We use the shorthand $\subs{\exprs{V}}{X^i}{X}$ to denote the formula
where every variable $x$ is substituted to $x^+$

We encode the finite set of cells $\Cells$ with a set of Boolean
variables $\bvars{}$. For clarity, given the cell $C \in \Cells$ we
write $\cellvar{} = C$, instead of showing its encoding with $\bvars{}$.
%
In the following, we interpret each variable $x \in \vx$ as a Real
valued variable and we refer to all the Boolean and Real variables
with the set $\vars{} = \rvars{} \vee \bvars{}$
%
We encode the set of paths of length $k$ of
$\RTSAbstraction{\epsilon}{\System}$ that reaches $\unsafebmc$
as the quantifier first-order logic formula:
\begin{eqnarray*}
\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k) & := &
\initbmc(\vars{0}) \wedge
\bigwedge_{i=1}^{k}{\transbmc(\vars{i-1}, \vars{i})} \wedge
\subs{\unsafebmc}{\rvars{k}}{\rvars{}}\\
% init
\initbmc(\vars{i}) & := &
\bigwedge_{C \in \InitCells}{(\cellvar{i} = C) \wedge
\subs{\CellLabelingFunction(C)}{\rvars{i}}{\rvars{}}
}
\\
% trans
\transbmc(\vars{i}, \vars{i+1}) & := &
  \bigvee_{(C,C^+) \in \CellLabelingFunction}{
    \edgebmc_{(C,C^+)}(\vars{i}, \vars{i+1})
  } \\
\edgebmc_{(C,C^+)}(\vars{i}, \vars{i+1}) & := &
(\cellvar{i} = C) \wedge (\cellvar{i+1} = C^+) \wedge  \\
&& \subs{\CellLabelingFunction(C)}{\rvars{i}}{\rvars{}} \wedge
   \subs{\CellLabelingFunction(C^+)}{\rvars{i+1}}{\rvars{}} \wedge \\
&& \subs{\subs{\EdgeLabelingFunction((C,C^+))}{\rvars{i}}{\rvars{}}}{\rvars{i+1}}{{\rvars{+}}}
\end{eqnarray*}

The formula
$\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$ is satisfiable
if and only if there exists a path of length $k$ in 
$\RTSAbstraction{\epsilon}{\System}$ that reaches a state in $\unsafebmc$.

The procedure
$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$
checks the satisfiability of the sequence of formulas 
$\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, 0)$, 
$\ldots$,
$\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$
with a SMT solver, returning as soon as one of the formula is
satisfiable, returning a path $\Pi$ of
$\RTSAbstraction{\epsilon}{\System}$.

\paragraph{Explicit path exploration.}
The second algorithm does not encode the reachability problem
$\mathsf{BoundedModelCheck}$ of a fixed length path as a single
satisfiability query, but it explicitly enumerates
the (abstract) paths in the graph that has $\Cells$
as nodes, and $\CellEdges$ as edges, and may end in an unsafe state
$\UnsafeStates$.
%
Such paths may not be real paths of
$\RTSAbstraction{\epsilon}{\System}$, since they may not satisfy the
conditions imposed by the labeling function $\CellLabelingFunction$
and the relation $\EdgeLabelingFunction$, and hence the algorithm must
further check, for each abstract path $\abspathbmc$, if tere exists a 
concrete path $\Pi$ of
$\RTSAbstraction{\epsilon}{\System}$.

$\abspathbmc := c_0, \ldots, c_l$ is an abstract path of
 $\RTSAbstraction{\epsilon}{\System}$ where:
\begin{itemize}[label=--,leftmargin=1em,labelsep=*]
\item for $i \in [0,l]$, $c_i \in \Cells$,
\item $c_0 \in \InitCells$,
\item for $i \in [0,l-1]$, $(c_i, c_{i+1}) \in \EdgeLabelingFunction$,
\item $c_l$ is such that
$\CellLabelingFunction(c_l) \cap \UnsafeStates \neq \emptyset$.
\end{itemize}
%
There exists a concrete path $\Pi$ of
$\RTSAbstraction{\epsilon}{\System}$ that traverses the sequences of
cells of an abstract path $\abspathbmc$ if and only if the formula
\begin{eqnarray}
\label{eq:simulateabs}
\concbmc{\abspathbmc} & := &
\bigg( \bigwedge_{i=0}^{l}{
  \subs{\CellLabelingFunction(c_i)}{\rvars{i}}{\rvars{}}
} \bigg)
\wedge
\subs{\unsafebmc}{\rvars{l}}{\rvars{}} \wedge \\
&&
\bigwedge_{i=1}^{l}{
  \subs{
  \subs{\CellEdges((c_i,c_{i+1}))}{\rvars{i}}{\rvars{}}
  }{\rvars{i+1}}{\rvars{+}}
}
\end{eqnarray}
is satisfiable\footnote{Formula~\ref{eq:simulateabs} is a conjunction
of linear predicates and hence its satisfiability can be casted as the
feasibility of a linear program}.
%
This can proved showing that each model to the formula
$\concbmc{\abspathbmc}$  ``corresponds'' to a path of
$\RTSAbstraction{\epsilon}{\System}$, and each path of
$\RTSAbstraction{\epsilon}{\System}$ that traverse the sequence
$\abspathbmc$ ``corresponds'' to a model of
Formula~\ref{eq:simulateabs}.

The explicit path exploration enumerates all the abstract
paths $\abspathbmc$ of length at most $k$. The enumeration can either
be a depth-first search (DFS) or a breadth-first search (BFS). As soon
as $\abspathbmc$ is found, the algorithm checks for the satisfiability
of $\concbmc{\abspathbmc}$. If  $\concbmc{\abspathbmc}$ is
satisfiable, the algorithm returns a concrete path $\Pi$, otherwise
the search continue trying to enumerate another abstract path. 
The search is guaranteed to terminate, since the length of the
abstract paths is at most $k$.

\todo{Tell that we use a DFS in the implementation section}

%% Temporal properties of PWA transitions systems can be checked using
%% off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
%% over infinite state transitions systems using SMT solvers.
%% Furthermore, lazy SMT solvers can be employed for this specific
%% problem instance to achieve better efficiency~\cite{shoukry2017smc}.
%% We now show how relational PWA abstractions can be viewed as PWA
%% transition systems, and how the problem of falsification can
%% be answered by a BMC query.
